Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-10 Manuscript/manuscript.tex
Update salt concentrations and binding affinity sections

While, the charge scaling improved the behaviour of lipid-ion interactions,
it reduced the area per molecule of the lipid bilayer without ions below experimental
values. Simulations with Lipid14 parameters having partial charges of head group, glycerol
backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$,
which is smaller than the experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
and the original Lipid14 value $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
The decrease of the area per lipid was found to arise from a lower hydration of the lipid head group region,
which can be explained by the increased solvation free energy due to the lower polarity
of molecules with scaled charges. The hydration can be increased back
by reducing the effective radius of atoms by changing the 
the $\sigma$ parameters in the Lennard-Jones potential for the selected atoms
similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules.
  JOE: I think that we rather increasingly see that there's nothing like ``fully non-scaled'' with the exception of ions with integer
  charges. So the discussion shall be rather more about the interaction of our ``scaled'' (I'd still rather call it ECC-corrected or whatever)
  and ``semi-scaled'' models.}
This decreases the solvation free energy by allowing water molecules to
approach closer to lipid atoms and have stronger electrostatic interactions with them.
After reducing the $\sigma$ parameters by a factor of $f_\sigma = 0.89$
for the same atoms for which charges were scaled, the area per molecule value was
back in agreement with the experimental value (see Table~\ref{tab:apls}).

@@

\subsection{Salt concentrations and binding affinity}
Experimental studies providing the lipid head group order
parameters report salt concentrations in two different ways.
For DPPC, salt concentrations before solvating
the lipids are used~\cite{akutsu81}. Alternatively, for POPC,
the salt concentration of supernatant after solvation is
measured using atomic absorption~\cite{altenbach84}.
\todo{JOE: I'm afraid that the above description might sound to the reader
  like for some reasons one method is used with DPPC, but a different one with POPC.
  I think simply writing one way is\dots another way is \dots (or similarly) will be less confusing.}
In this work, we will use the latter definition. For this we
will calculate the salt concentration in the aqueous bulk region,
\textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
Note that in the previous study, Ref.~\citenum{catte16},
the concentration was calculated as in the DPPC experiments.
Despite of the measurable differences between these two concentrations for CaCl$_2$ systems,
the qualitative conclusions are affected neither in this
nor in our previous work~\cite{catte16}.

To quantify the ion binding affinity to a membrane, we
calculated the relative surface excess of ions with respect to water $\Gamma_i^w$  \cite{chattorajBOOK}.
This quantity can be used to compare binding affinity between different models,
because it does not depend on the definition of the interfacial region
or the position of the Gibbs dividing plane between two bulk regions.
\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
Therefore, we consider the hydrophobic interior of a membrane having zero concentrations of ions and water
as a bulk region, which is separated by a monolayer interface from 
a another region with bulk concentration of ions in water.
Since the boundary of the latter region is set to the edge of simulation box,
all molecules in simulation are within the defined interfacial region. 
%our whole simulation box as
%an interface flanked by a water phase and an ideal phase with the characteristics of to
%the hydrophobic region in the middle of the membrane where no ions or waters are present.
This setup provides a simplified relation for $\Gamma_i^w$ in lipid bilayers simulations 
\begin{equation}\label{surfexcess}
  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
\end{equation}
where $n_w$ and $n_i$ are the total number of waters and ions in the system;
$C_w$ and $C_i$ are their respective bulk concentrations in the aqueous phase;
and $A_b$ is the size of the box in the membrane plane.
The total area of the interface is twice the area of the membrane, $2A_b$,
because bilayers have an interface at both leaflets.

>>>>

While, the charge scaling improved the behaviour of lipid-ion interactions,
it reduced the area per molecule of the lipid bilayer without ions below experimental
values. Simulations with Lipid14 parameters having partial charges of head group, glycerol
backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$,
which is smaller than the experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
and the original Lipid14 value $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
The decrease of the area per lipid was found to arise from a lower hydration of the lipid head group region,
which can be explained by the increased solvation free energy due to the lower polarity
of molecules with scaled charges. The hydration can be increased back
by reducing the effective radius of atoms by changing the 
the $\sigma$ parameters in the Lennard-Jones potential for the selected atoms
similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules. \\
  JOE: I think that we rather increasingly see that there's nothing like ``fully non-scaled'' with the exception of ions with integer
  charges. So the discussion shall be rather more about the interaction of our ``scaled'' (I'd still rather call it ECC-corrected or whatever)
  and ``semi-scaled'' models. \\
SAMULI: There is now a paragraph in the conclusions, which mentions this topic.}
This decreases the solvation free energy by allowing water molecules to
approach closer to lipid atoms and have stronger electrostatic interactions with them.
After reducing the $\sigma$ parameters by a factor of $f_\sigma = 0.89$
for the same atoms for which charges were scaled, the area per molecule value was
back in agreement with the experimental value (see Table~\ref{tab:apls}).

@@

\subsection{Salt concentrations and binding affinity}
The early experimental electrometer concept data for the lipid head group
order parameter changes was reported as a function 
of the salt concentration in water before solvating
the lipids~\cite{akutsu81}. The later study used atomic absorption
spectroscopy and reported the order parameter changes as
a function of the salt concentration in the supernatant after the
solvation of lipids~\cite{altenbach84}.
In this work we focus on POPC for which the latter definition was used~\cite{altenbach84}.
The salt concentration in the aqueous bulk region was calculated from 
the farthest point from both lipid leaflets in the water phase. 
Note that in the previous study, Ref.~\citenum{catte16},
the ion concentrations were calculated in water before solvating the lipids 
as in the earlier experiments~\cite{akutsu81}.
Despite of the measurable differences between these two concentrations for CaCl$_2$ systems,
the qualitative conclusions in this or in the previous work~\cite{catte16} are
not affected.

To quantify the ion binding affinity to a lipid bilayer, we
calculated the relative surface excess of ions with respect to water,~$\Gamma_i^w$~\cite{chattorajBOOK}.
This quantity was chosen because it does not depend on
%the definition of the interfacial region or
the position of the Gibbs dividing plane between two bulk regions.
%\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
%SAMULI: Here is a new trial.
Here we assume that the interface locates between the hydrophobic interior
of a lipid bilayer and the bulk water region far from the membrane.
The bulk concentration of ions and water is zero inside the bilayer.
The concentrations in bulk water region can be calculated from 
the farthest point from both lipid leaflets in the water phase.
The region between these boundaries contains all the ions and water
molecules in the simulation box.
%our whole simulation box as
%an interface flanked by a water phase and an ideal phase with the characteristics of to
%the hydrophobic region in the middle of the membrane where no ions or waters are present.
This setup provides a simplified relation for $\Gamma_i^w$ in lipid bilayers simulations 
\begin{equation}\label{surfexcess}
  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
\end{equation}
where $n_w$ and $n_i$ are the total number of waters and ions in the system;
$C_w$ and $C_i$ are their respective bulk concentrations in the aqueous phase;
and $A_b$ is the size of the box in the membrane plane.
The total area of the interface is twice the area of the membrane, $2A_b$,
because bilayers have an interface at both leaflets.
